Issues with Dynamical Measurements Across the Disk

### 2019-02-12 paper/ms.tex
add selection function statement
revisions to sections 2.1 and 2.2

We do not propose what a modeller should use for the ``true'' global midplane,
since such a definition probably depends on the particular problem such a
modeller is facing. Even without a high-precision definition and measurement
of the global Galactic midplane, we can still assess how different we should
expect the local midplane to be by measuring the azimuthal dependence of the
local midplane. This procedure may be readily tractable over some of the disk
with current datasets. However, we turn to simulation in this work because we
can investigate the entire disk. Specifically, we use cosmological zoom-ins of
Milky Way-mass galaxies from the Feedback in Realistic Environments (FIRE)
collaboration\footnote{\url{https://fire.northwestern.edu}} (the {\em Latte}
suite of simulations).

@@

\subsection{Cartoon} \label{ssec:cartoon}
We present a cartoon in Figure~\ref{fig:cartoon} to show that an inaccurate
determination of the midplane can lead to a phase-dependence of the {\em
observed} actions. The solid gray line indicates a hypothetical ``true''
orbit. The y-axis corresponds to the vertical height of the orbit and the
x-axis orbital phase. The dashed gray line is an incorrect midplane which an
observer uses to compute orbits.

Now suppose this observer makes a measurement of the true orbit at the blue
point or the pink point. Then the blue and pink lines correspond to the orbits
that the observer would compute for each point. In action space, this would
correspond to a different value of $J_z$ for the blue and pink points. In this
way, assuming the wrong coordinate system induces time dependence in the
actions {\em computed by an observer}.

\begin{figure*}
\plotone{fig/cartoon.pdf}
\caption{Cartoon showing the effect an incorrect midplane can have on orbit
integrations and as an extension on action computations. The x-axis shows the
orbital phase and the y-axis the vertical height. The top gray curve depicts a
hypothetical true orbit oscillating about the true midplane, the horizontal
solid gray line. Consider an immortal observer who erroneously assumes the
midplane is located at the horizontal dashed line. Suppose this immortal
observer measures the orbit at the blue and pink points. If the immortal
observer were to perform orbit integrations at each observation, they would
assume the blue and pink curves for the respective orbits. In this way, an
incorrect midplane will induce phase-dependence in the {\em observed}
actions.}
\label{fig:cartoon}
\end{figure*}

\subsection{Summary of Action Computation} \label{ssec:action_comp}
We follow our previous work in computing actions \citep{2018ApJ...867...31B}.
We use the code \texttt{gala} v0.3 to perform orbit integrations and
conversion to action space \citep{2017JOSS....2..388P,Price-Whelan:2018},
which uses an action estimator technique presented by
\citet{2014MNRAS.441.3284S}. We use the default \texttt{MWPotential} as our
potential, which is based on the Milky Way potential available in
\texttt{galpy} \citep{2015ApJS..216...29B}. This potential includes a
Hernquist bulge and nucleus \citep{1990ApJ...356..359H}, a Miyamoto-Nagai disk
\citep{1975PASJ...27..533M}, and an NFW halo \citep{1997ApJ...490..493N}, and
is fit to empirically match some observations. We use the Dormand-Prince
8(5,3) integration scheme \citep{Dormand80:integrator}. We use a timestep of
$1\,\Myr$ and integrate for $5\,\Gyr$, corresponding to $\sim 20$ orbits for a
Sun-like star.

Other methods for computing actions are used in the literature. For example,
the St\"ackel Fudge method \citep{2016MNRAS.457.2107S} which uses the
St\"ackel potential to approximate the Galactic potential
\citep{1985MNRAS.216..273D,2012MNRAS.426.1324B} is commonly used
\citep{2018arXiv180503653T,2018MNRAS.481.4093S,2018arXiv180803278T}. For
applications in the disk, the St\"ackel Fudge method is of acceptable
accuracy, but since we also consider halo orbits in this work we choose to use
orbit integration throughout \citep{2016MNRAS.457.2107S}.

>>>>

We do not propose what a modeller should use for the ``true'' global midplane,
since such a definition probably depends on the particular problem such a
modeller is facing. Even without a high-precision definition and measurement
of the global Galactic midplane, we can still assess how different we should
expect the local midplane to be by measuring the azimuthal dependence of the
local midplane. This procedure may be readily tractable over some of the disk
with current datasets, using appropriate selection functions. However, we turn
to simulation in this work because we can investigate the entire disk.
Specifically, we use cosmological zoom-ins of Milky Way-mass galaxies from the
Feedback in Realistic Environments (FIRE)
collaboration\footnote{\url{https://fire.northwestern.edu}} (the {\em Latte}
suite of simulations).

@@

\subsection{Cartoon} \label{ssec:cartoon}
We present a cartoon in Figure~\ref{fig:cartoon} to show that an inaccurate
determination of the midplane can lead to a phase-dependence of the actions calculated from an orbit specified by a phase-space point and an assumed potential. The y-axis corresponds to the vertical height of the orbit and the x-axis orbital phase. The solid gray line indicates the ``true'' orbit of the star; that is, its orbit as it oscillates around the true midplane. The dashed gray line is an erroneous midplane assigned to the potential model which an observer uses to integrate the orbit of the star and hence estimate its actions. The model potential is otherwise identical to the one in which the star is actually moving.

Now suppose this observer makes a measurement of the true orbit at the blue
point or the pink point (i.e. at two different orbital phases). Then the blue and pink lines correspond to the orbits that the observer would compute for each point based on the potential model with the erroneous midplane. In action space, this would correspond to a different value of $J_z$ for the blue and pink points. In this way, assuming the wrong coordinate system induces a phase-dependence in the
actions estimated for the star, which in the correct potential (in this example, the one with the correct midplane) should be phase-independent.

\begin{figure*}
\plotone{fig/cartoon.pdf}
\caption{Cartoon showing the effect an error in the determination of the coordinate midplane can have on orbit integration and action estimation. The x-axis shows the
orbital phase and the y-axis the vertical height. The top gray curve depicts an example ``true'' orbit oscillating about the true midplane (horizontal
solid gray line). Consider an observer who erroneously assumes the
midplane is located at the horizontal dashed line. Suppose this
observer measures the phase-space position of the star at two different orbital phases (blue and pink points). If the observer were then to integrate the star's orbit using a model potential with the erroneous midplane, they would
obtain the blue and pink curves for the star's orbit, respectively. The actions estimated from these two  erroneous orbits would subsequently differ, both from each other and from the actions estimated for the true orbit (in the potential with the correct midplane). Hence an
incorrect midplane in the potential model assumed will induce phase-dependence in the actions estimated for a given star in that potential.}
\label{fig:cartoon}
\end{figure*}

\subsection{Summary of Action Computation} \label{ssec:action_comp}
We compute actions as in our previous work \citep{2018ApJ...867...31B}, using the code \texttt{gala} v0.3 to perform orbit integrations and
conversion to action space \citep{2017JOSS....2..388P,Price-Whelan:2018}.
To compute actions we use the torus-mapping technique first presented by \citet{1990MNRAS.244..634M} and adapted by \citet{2014MNRAS.441.3284S} to calculate actions for an orbital time-series starting from a phase-space position $(x, v)$ and integrated in a potential $\Phi$. We use the default \texttt{MWPotential} as our
potential, which is based on the Milky Way potential available in
\texttt{galpy} \citep{2015ApJS..216...29B}. This potential includes a
Hernquist bulge and nucleus \citep{1990ApJ...356..359H}, a Miyamoto-Nagai disk
\citep{1975PASJ...27..533M}, and an NFW halo \citep{1997ApJ...490..493N}, and
is fit to empirically match some observations. We use the Dormand-Prince
8(5,3) integration scheme \citep{Dormand80:integrator} with a timestep of
$1\,\Myr$ and integrate for $5\,\Gyr$, corresponding to $\sim 20$ orbits for a
Sun-like star.

Other methods for computing actions are used in the literature. For example,
the St\"ackel Fudge method \citep{2016MNRAS.457.2107S}, which uses a single
St\"ackel potential (with analytic actions) to approximate the Galactic potential
\citep{1985MNRAS.216..273D,2012MNRAS.426.1324B}, was used in many recent works exploring actions in the Galactic disk
\citep[e.g.][]{2018arXiv180503653T,2018MNRAS.481.4093S,2018arXiv180803278T}. For
disk-like orbits, existing implementations of the St\"ackel Fudge method are of acceptable
accuracy, but since we also consider halo-like orbits in this work we choose to use
orbit integration and torus mapping throughout \citep{2016MNRAS.457.2107S}.
